# Advanced R - Exercises and notes

_Links:_

[Advanced R](http://adv-r.had.co.nz)

[UBC Discussion on Github](https://github.com/aammd/UBCadv-r)

[UBC Discussion on Github Wiki](https://github.com/aammd/UBCadv-r/wiki)

[R fiddle](http://www.r-fiddle.org/)

---
__Table of Content__

[TOC]

---

##Subsetting 

### Data types
#####Exercise
(1) Fix each of the following common data frame subsetting errors:
```{r}
mtcars[mtcars$cyl = 4, ]
mtcars[-1:4, ]
mtcars[mtcars$cyl <= 5]
mtcars[mtcars$cyl == 4 | 6, ]
```
```{r}
mtcars[mtcars$cyl == 4, ]
mtcars[-c(1:4),]
mtcars[mtcars$cyl <= 5, ]
mtcars[mtcars$cyl %in% c(4,6), ]
```

(2) Why does x <- 1:5; x[NA] yield five missing values? (Hint: why is it different from x[NA_real_]?)

> `NA`is a logical vector, and logical vector is valid for indexing. 
>  Indexed by a logical vector, it compares each element and return if `TRUE`. 

```{r}
x = 1:5 # [1] 1 2 3 4 5
x[TRUE] # [1] 1 2 3 4 5
x[NA] # [1] NA NA NA NA NA
```

(3) What does upper.tri() return? How does subsetting a matrix with it work? Do we need any additional subsetting rules to describe its behaviour?
```{r}
x <- outer(1:5, 1:5, FUN = "*")
x[upper.tri(x)]
```
```{r}
#> x
#     [,1] [,2] [,3] [,4] [,5]
#[1,]    1    2    3    4    5
#[2,]    2    4    6    8   10
#[3,]    3    6    9   12   15
#[4,]    4    8   12   16   20
#[5,]    5   10   15   20   25

#> upper.tri(x)
#      [,1]  [,2]  [,3]  [,4]  [,5]
#[1,] FALSE  TRUE  TRUE  TRUE  TRUE
#[2,] FALSE FALSE  TRUE  TRUE  TRUE
#[3,] FALSE FALSE FALSE  TRUE  TRUE
#[4,] FALSE FALSE FALSE FALSE  TRUE
#[5,] FALSE FALSE FALSE FALSE FALSE

# the subsetting then return the upper tri element in the form of a single line vector.
```
 (4) Why does mtcars[1:20] return a error? How does it differ from the similar mtcars[1:20, ]?

> It returns error since `mtcars` is 2 dimensions data. 
> `mtcars[1:20, ]` works as it involve the 2nd dimension.

(5) Implement your own function that extracts the diagonal entries from a matrix (it should behave like diag(x) where x is a matrix).

```{r}
diag(x) # [1]  1  4  9 16 25

## Method 1 
x = outer(1:5, 1:5, FUN = "*")
diag_x = function (x) {
    d=c()
    for (i in 1:nrow(x)){
        d[i] = x[i,i]
    }
    return (d)
}
diag_x (x) # [1]  1  4  9 16 25

## Method 2
diag_x2 = function (x) {
    y = seq(1, nrow(x)^2, nrow(x)+1)
    return (x[y])
}
diag_x2(x) # [1]  1  4  9 16 25
```

(6) What does df[is.na(df)] <- 0 do? How does it work?

> It will return a logical vector with whether each element of `df` is `NA`. If `TRUE`, assign 0 into it.

### Subsetting operators
#####Exercise
(1) Given a linear model, e.g., mod <- lm(mpg ~ wt, data = mtcars), extract the residual degrees of freedom. Extract the R squared from the model summary (summary(mod))

``` {r}
mod <- lm(mpg ~ wt, data = mtcars)

# Extract the residual degree of freedom
names(mod) # get the name of the list
mod[["residuals"]]

# Extract the R squared from the model summary
summaryMod = summary(mod)
names(summaryMod)
summaryMod[["r.squared"]]
```

### Subsetting and assignment

####Matching and merging by hand
#####Note
*  Matching lookup table
```{r}
# Data
grades <- c(1, 2, 2, 3, 1)

info <- data.frame(
  grade = 3:1,
  desc = c("Excellent", "Good", "Poor"),
  fail = c(F, F, T)
)
```
```{r}
# match lookup table
grades
#> [1] 1 2 2 3 1

# Using match
id <- match(grades, info$grade)
info[id, ]
#>     grade      desc  fail
#> 3       1      Poor  TRUE
#> 2       2      Good FALSE
#> 2.1     2      Good FALSE
#> 1       3 Excellent FALSE
#> 3.1     1      Poor  TRUE

# Using rownames
rownames(info) <- info$grade
info[as.character(grades), ]
#>     grade      desc  fail
#> 1       1      Poor  TRUE
#> 2       2      Good FALSE
#> 2.1     2      Good FALSE
#> 3       3 Excellent FALSE
#> 1.1     1      Poor  TRUE
```

* match multiple columns
```{r}
data = data.frame(
    grade = c(1, 2, 2, 3, 1),
    type = c("A", "B","A", "B","A")
)

info <- data.frame(
  type = c(rep("A",2),rep("B",3)),
  grade = c(2,1,3,2,1),
  desc = c("Pass","Fail","Excellent", "Good", "Poor"),
  fail = c(F, T,F, F, T)
)

# use interaction() to collapse keys to single column
info$key = interaction(info[,c(2,1)])

data_key = interaction(data)

id = match(data_key, info$key)
info[id,]

# merge()
merge_data = merge(data, info, by=c("type", "grade"))
merge_data
```

(1) How would you randomly permute the columns of a data frame? (This is an important technique in random forests.) Can you simultaneously permute the rows and columns in one step?
```{r}
df <- data.frame(x = rep(1:3, each = 2), y = 6:1, z = letters[1:6])

# permute columns
df_col = df[,sample(1:ncol(df))]
df_col

# permute rows and columns
df_rc = df[sample(1:nrow(df)), sample(1:ncol(df))]
df_rc

```

(2) How would you select a random sample of m rows from a data frame? What if the sample had to be contiguous (i.e., with an initial row, a final row, and every row in between)?
```{r}
df <- data.frame(x = rep(1:3, each = 2), y = 6:1, z = letters[1:6])
m = 2

# select m random rows
df1 = df[sample(1:nrow(df),m),]
df1

# select m countinous rows
k = sample(1:(nrow(df)-m+1),1)
df2 = df[k:(k+m-1),]
df2
```

(3) How could you put the columns in a data frame in alphabetical order?
```{r}
df <- data.frame(x = rep(1:3, each = 2), y = 6:1, z = letters[1:6])

# randomly reorder df
df1 = df[, sample(1:ncol(df))]
df1

# sort columns in alphabetical order
df2 = df[, sort(names(df1))]
# OR 
df2 = df[, order(names(df1))]
df2
```

##Functions
###Function components
(1) What function allows you to tell if an object is a function? What function allows you to tell if a function is a primitive function?

```{r}
is.function(sum) # [1] TRUE
is.primitive(sum) # [1] TRUE
```

(2) This code makes a list of all functions in the base package.

```{r}
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)
```
Use it to answer the following questions:
a. Which base function has the most arguments?
b. How many base functions have no arguments? What’s special about those functions?
c. How could you adapt the code to find all primitive functions?

```{r}
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)

## a
arg_length = lapply(funs, function(x)(length(formals(x))))
which.max(arg_length) # scan  933 

## b
arg_0 = which(arg_length == 0)
length(arg_0) # 221
names(arg_0) 

## c
arg_prim = lapply(funs, is.primitive)
which(arg_prim==T) 
```

(3) What are the three important components of a function?
> _from the content_
> * __body__: the code inside the function.
> * __formals__ :  the list of arguments which controls how you can call the function.
> * __environment__ : the “map” of the location of the function’s variables.

(4) When does printing a function not show what environment it was created in?
> _If the environment isn’t displayed, it means that the function was created in the global environment._

###Lexical scoping
(1) What does the following code return? Why? What does each of the three c’s mean?
```{r}
c <- 10
c(c = c)
```
> return 10 as a list
> 1st c : function
> 2nd c: argumnet
> 3rd c: variable

(2) What are the four principles that govern how R looks for values?
> * name masking
> * functions vs. variables
> * a fresh start
> * dynamic lookup

(3) What does the following function return? Make a prediction before running the code yourself.
```{r}
f <- function(x) {
  f <- function(x) {
    f <- function(x) {
      x ^ 2
    }
    f(x) + 1
  }
  f(x) * 2
}
f(10)
```
> (((10)^2 )+1 )*2 = 202

#####note
Calling a function given a list of arguments 
```{r} 
do.call(mean, list(1:10, na.rm = TRUE))
# <=>
mean(1:10, na.rm = TRUE)
```

####Function arguments
(1) Clarify the following list of odd function calls:
```{r}
x <- sample(replace = TRUE, 20, x = c(1:10, NA))
y <- runif(min = 0, max = 1, 20)
cor(m = "k", y = y, u = "p", x = x)
```
```{r}
x = sample(x = c(1:10, NA), size = 20, replace = TRUE)
y = runif(20, min = 0, max = 1)
cor(x, y, use = "pairwise", method = "kendall")
```

(2) What does this function return? Why? Which principle does it illustrate?
```{r}
f1 <- function(x = {y <- 1; 2}, y = 0) {
  x + y
}
f1()
```
```{r}
x + y -> (y=1, x=2) + y -> 2 + 1 = 3
```

(3) What does this function return? Why? Which principle does it illustrate?
```{r}
f2 <- function(x = z) {
  z <- 100
  x
}
f2()
```
```{r}
# 1 
z = 100
# 2 
x = z (100)
```
